An efficient acyclic contact planner for multiped robots

### 2017-01-10 body/intro.tex
accepted andrea s update on introduction

%~ Planning legged robot locomotion remains an open issue.
Legged robots move by sequentially creating contacts with the environment\deladp{ while avoiding collisions}.
After years of research, such robots \addadp{can} autonomously walk on flat ground, but struggle to \repadp{navigate more complex environments}{achieve more complex motions}.
To climb stairs using a handrail for instance, a humanoid robot must plan the contacts to create with its feet and hands \deladp{to perform the motion safely}. \deladp{The issue of} computing \repadp{such a}{this} contact sequence is a\addadp{n open} motion planning problem called acyclic contact planning.

Once the contact plan is known, efficient approaches \repadp{exist to}{can execute the plan and} generate a dynamically feasible motion~\citep{Carpentier2016}. However in general
the contact plan is manually designed \repadp{by a time consuming trial-and-error procedure}{after long trials and error}. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in human-centered environments, which are cluttered and comprise constraining obstacles such as stairs.

To make a legged robot walk on \deladp{a} flat ground, efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. Furthermore, choosing the effector with which to create a contact is trivial because walk\addadp{ing} follows a cyclic pattern (For a humanoid the left foot always follows the right foot).
Unfortunately, in the general case planning complex contact interactions is extremely challenging:
At any given time a contact must be chosen between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles; we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. Lastly, the contact manifold has a null measure and is thus impossible to sample. This restriction prevents the use of efficient sampling-based planners\deladp{ to solve the problem}.

@@

The issue of planning acyclic contacts was first completely described by \citeauthor{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two sub-problems, $\mathcal{P}_1$: planning a \deladp{relevant} guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{Bretl:2006:MPM:1124573.1124585} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the \repadp{method}{computation time} is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. \citeauthor{DBLP:conf/humanoids/DeitsT14} proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered. 
\citeauthor{dai2014whole} extended the work of \citeauthor{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

@@

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for \deladp{the motion of} a humanoid character, and managed to obtain dynamically-consistent motions
at \gls{interactive} frame rates (we say that a method is interactive when the computation time for planning is less than the
time for execution). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold. \deladp{The space is sampled thanks to the \textit{reachability condition}, a low-dimensional geometric condition for \textit{contact reachability}.}

@@

Finally, to further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times \deladp{we obtained} to address sequentially the three problems, and compare them \deladp{with respect} to related work\deladp{, thus clearly} demonstrating that our method is order of magnitude faster\deladp{ than any existing method}.

>>>>

%~ Planning legged robot locomotion remains an open issue.
Legged robots move by sequentially creating contacts with the environment.
After years of research, such robots can autonomously walk on flat ground, but struggle to navigate more complex environments.
To climb stairs using a handrail, a humanoid robot must plan the contacts to create with its feet and hands. Computing such a contact sequence is an open motion planning problem called acyclic contact planning.

Once the contact plan is known, efficient approaches exist to generate a dynamically feasible motion~\citep{Carpentier2016}. However in general
the contact plan is manually designed by a time consuming trial-and-error procedure. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in human-centered environments, which are cluttered and comprise constraining obstacles such as stairs.

To make a legged robot walk on flat ground, efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. Furthermore, choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (For a humanoid the left foot always follows the right foot).
Unfortunately, in the general case planning complex contact interactions is extremely challenging:
At any given time a contact must be chosen between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles; we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. Lastly, the contact manifold has a null measure and is thus impossible to sample. This restriction prevents the use of efficient sampling-based planners.

@@

The issue of planning acyclic contacts was first completely described by \citeauthor{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two sub-problems, $\mathcal{P}_1$: planning a guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{Bretl:2006:MPM:1124573.1124585} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. \citeauthor{DBLP:conf/humanoids/DeitsT14} proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered. 
\citeauthor{dai2014whole} extended the work of \citeauthor{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

@@

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
at \gls{interactive} frame rates (we say that a method is interactive when the computation time for planning is less than the
time for execution). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

@@

Finally, to further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is order of magnitude faster.
